HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-09 paper/main.tex
bbb

Human judgement is the central, and perhaps ultimate, criterion for estimating the
quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence, as is reflected
by the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated in the output, and hence cannot inform developers in overcoming these errors.
This problem can be addressed by using scores that decompose over the sub-parts of the sentence. 
However, common decompositions into n-grams or even individual words, as employed by the
popular BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005},
are most suitable for evaluating shallow MT approaches and are
less instructive for more structurally-aware methods, which are at
the forefront of MT research today.

These difficulties have motivated the definition of a human measure that decomposes over
the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
this type is the MEANT measure, and its human variant HMEANT
\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
similarity between the output translation and a reference translation in terms of the similarity
between their verbal argument structures
While the measure addresses the two major shortcomings
detailed above, it is unable to capture semantic differences which do not show up in
the verb-argument structure, including several very frequent phenomena, such as
inter-clause linkage, nominalizations and copula clauses (see \secref{sec:hmeant_comp}).
Another shortcoming of MEANT is its reliance on semantically annotating MT output,
which is often unintelligible. Finally, measures based on a comparison to a reference translation
are necessarily incomplete, as they compare to a small number of references
out of the the huge number of possible translations a sentence may have \cite{dryer_marcu:12}.

In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the sentence's semantic units.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

HUME operates by aggregating human assessments of the translation quality of the different
semantic units in the source sentence.
In thus only requires the semantic annotation of the source sentence, thus avoiding the 
annotatation of machine-generated text, a difficult if not ill-defined task.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
See \secref{sec:} for a discussion of HUME's advantages realative to other measures.

We conduct experiments with four language pairs, English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of resulting
HUME scores and BLEU, and crowd-sourced adequacy judgements. Results show that HUME obtains 
agreements which are in par with standard sentence-level ranking measures, but unlike with them,
the agreement does not diminish with the sentence length.

While keeping a human in the loop is never a scalable solution, human evaluation is the
fundamental measure of quality of MT output, and we therefore begin our inquiries here.

\oa{summary of contribution}

@@

>>>>

Human judgement is the central, and perhaps ultimate, criterion for estimating the
quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence, as is reflected
by the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated in the output, and hence cannot inform developers in overcoming these errors.

These problems were partially addressed by measures that decompose over sub-parts of the sentence,
most commonly over its words or n-grams, quantifying the overlap of the sub-parts with a reference
translation. More recent work on MT evaluation, most notably the MEANT measure \cite{lo2011structured},
proposed measures that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of their verb argument structures.
While allowing a localization of the semantic errors in the translation,
focusing on verb argument structures misses out on many very pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus provides only a limited
perspective on the semantic similarity between the output and reference.
Moreover, assigning a semantic structure to MT output, which is often low in quality,
is difficult and arguably ill-defined.

In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires the semantic annotation of the source sentence, thus avoiding the 
annotatation of machine-generated text, a difficult if not ill-defined task, and allowing the re-use
of the source semantic annotation for measuring the quality of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
See \secref{sec:hmeant_comp} for further discussion.

We conduct experiments with four language pairs, English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreements which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's agreement does not diminish with the sentence length.

While human evaluation is not a scalable solution,
it provides the foundations and justification for defining a measure.
The paper thus focuses on human evaluation, but concludes
by outlining ongoing work to construct semi- and fully-automatic variants of HUME.

@@

While the measure addresses the two major shortcomings
detailed above, it is unable to capture semantic differences which do not show up in
the verb-argument structure, including several very frequent phenomena, 
Another shortcoming of MEANT is its reliance on semantically annotating MT output,
which is often unintelligible. Finally, measures based on a comparison to a

This problem can be addressed by using scores that decompose over the sub-parts of the sentence. 
However, common decompositions into n-grams or even individual words, as employed by the
popular BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005},
are most suitable for evaluating shallow MT approaches and are
less instructive for more structurally-aware methods, which are at
the forefront of MT research today.

These difficulties have motivated the definition of a human measure that decomposes over
the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
this type is the MEANT measure, and its human variant HMEANT
\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
similarity between the output translation and a reference translation in terms of the similarity
between their verbal argument structures
While the measure addresses the two major shortcomings
detailed above, it is unable to capture semantic differences which do not show up in
the verb-argument structure, including several very frequent phenomena, such as
inter-clause linkage, nominalizations and copula clauses
(see \secref{sec:hmeant_comp}).
Another shortcoming of MEANT is its reliance on semantically annotating MT output,
which is often unintelligible. Finally, measures based on a comparison to a
reference translation
are necessarily incomplete, as they compare to a small number of references
out of the the huge number of possible translations a sentence may have
\cite{dreyer2012hyter}.
